---
title: "New Olink Study"
author: "Zaynab Mousavian"
date: "23 July 2024"
output:
  html_document: default
  pdf_document: default
subtitle: Support vector machine
---

```{r setup, include=FALSE}
# Set global options for all chunks
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Load R packages

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(multiview)
library(tidyr)
library(tidyverse)
library(caret)
library(pROC)
library(pROC)
library(caret)
library(groupdata2)
library(kernlab)
library(Hmisc)
library(tidyverse)
library(ROCaggregator)
library(RColorBrewer)
library(ROCR)
library(pracma)
library(stats)
library(caretEnsemble)
library(PRROC)
```

# Function for visualization ROC plot with Youden index with text

```{r}
ROCplotwithYoudenText_func <- function(rocObj, ciData, specific_specificity = NULL, marker_name=NULL){
  
  youden_index <- rocObj$sensitivities + rocObj$specificities - 1
  optimal_threshold_index <- which.max(youden_index)
  optimal_threshold <- rocObj$thresholds[optimal_threshold_index]

  optimal_sensitivity <- rocObj$sensitivities[optimal_threshold_index]
  optimal_specificity <- rocObj$specificities[optimal_threshold_index]
 
  plot <- ggroc(rocObj,legacy.axes = TRUE)  + geom_abline(slope=1, intercept = 0, linetype = "dashed", alpha=0.7, color = "grey") + coord_equal() + 
    geom_ribbon(data = ciData, aes(x = 1-x, ymin = lower, ymax = upper), fill = "steelblue", alpha= 0.2) + 
    ggtitle(paste0("AUC=",round(roc0$auc,2)," (",gsub(" \\(DeLong\\)","",capture.output(roc0$ci)),")"))+
    labs(x = "1 - Specificity",
         y = "Sensitivity")+
    theme(axis.text.x = element_text(size = 12),  
          axis.text.y = element_text(size = 12),
          axis.title.y = element_text(size=12),
          plot.title = element_text(size=9),
          panel.background = element_rect(fill = "white", color = NA),  
          plot.background = element_rect(fill = "white", color = NA), 
          panel.grid.major = element_blank(),  
          panel.grid.minor = element_blank(),
          panel.border = element_rect(color ="grey", fill = NA))+
  
    geom_point(aes(x = 1 - optimal_specificity, y = optimal_sensitivity), color = "black", size = 2) +
    geom_text(aes(x = 1 - optimal_specificity, y = optimal_sensitivity, 
                  label = paste0("(%", round(optimal_sensitivity, 2)*100,", %",round(optimal_specificity, 2)*100, ")")), 
              hjust = 1, vjust = 1.5, x = 0.95, y = 0.05,color = "black", size = 4)
    if (!is.null(specific_specificity)) {
    specific_fpr <- 1 - specific_specificity
    
    
    closest_specificity_index <- which.min(abs(rocObj$specificities - specific_specificity))
    sensitivity_at_specific_specificity <- rocObj$sensitivities[closest_specificity_index]
    
   
    plot <- plot +
      geom_point(aes(x = specific_fpr, y = sensitivity_at_specific_specificity), color = "lightcoral", size = 2) +
      geom_text(aes(x = specific_fpr, y = sensitivity_at_specific_specificity, 
                    label = paste0("(%", round(sensitivity_at_specific_specificity, 2)*100,", %",round(specific_specificity, 2)*100, ")")), 
                hjust = 1, vjust = -0.5,x = 0.95, y = 0.05, color = "lightcoral", size = 4)
    }
  if (!is.null(marker_name)) {
  
  plot <- plot +
    annotate("text", x = 0.95, y = 0.05, 
             label = marker_name, 
             hjust = 1, vjust = 0, 
             color = "blue", size = 4, fontface = "italic")
  }
  return(plot)
}
```



# Function for visualization ROC plot with Youden index

```{r}
ROCplotwithYouden_func <- function(rocObj, ciData, specific_specificity = NULL,marker_name=NULL){
  
  youden_index <- rocObj$sensitivities + rocObj$specificities - 1
  optimal_threshold_index <- which.max(youden_index)
  optimal_threshold <- rocObj$thresholds[optimal_threshold_index]

  optimal_sensitivity <- rocObj$sensitivities[optimal_threshold_index]
  optimal_specificity <- rocObj$specificities[optimal_threshold_index]
 
  plot <- ggroc(rocObj,legacy.axes = TRUE)  + geom_abline(slope=1, intercept = 0, linetype = "dashed", alpha=0.7, color = "grey") + coord_equal() + 
    geom_ribbon(data = ciData, aes(x = 1-x, ymin = lower, ymax = upper), fill = "steelblue", alpha= 0.2) + 
    ggtitle(paste0("AUC=",round(roc0$auc,2)," (",gsub(" \\(DeLong\\)","",capture.output(roc0$ci)),")"))+
    labs(x = "1 - Specificity",
         y = "Sensitivity")+
    theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12),
          axis.title.y = element_text(size=12),
          plot.title = element_text(size=9),
          panel.background = element_rect(fill = "white", color = NA),  
          plot.background = element_rect(fill = "white", color = NA), 
          panel.grid.major = element_blank(),  
          panel.grid.minor = element_blank(),
          panel.border = element_rect(color = "gray", fill = NA))+
    geom_point(aes(x = 1 - optimal_specificity, y = optimal_sensitivity), color = "black", size = 2) 
 if (!is.null(specific_specificity)) {
    specific_fpr <- 1 - specific_specificity
    closest_specificity_index <- which.min(abs(rocObj$specificities - specific_specificity))
    sensitivity_at_specific_specificity <- rocObj$sensitivities[closest_specificity_index]
    
    plot <- plot + 
      geom_point(aes(x = specific_fpr, y = sensitivity_at_specific_specificity), color = "lightcoral", size = 2) 
 }
   if (!is.null(marker_name)) {
  
  plot <- plot +
    annotate("text", x = 0.95, y = 0.05, 
             label = marker_name, 
             hjust = 1, vjust = 0, 
             color = "blue", size = 4, fontface = "italic")
  }
  return(plot)
}
```


# Which data? {.tabset}

## Signature including 12 proteins

### Load data for TB and nonTB from Italy

```{r message=FALSE, warning=FALSE}
l=load("../Data/ReadyData/AllDatasets-Samples.RData")
load(file="../Data/ReadyData/generated for iScience/0.3-pearson-modules.RData")
module="turquoise"
column = match(module, modNames);
moduleGenes = modules [[column]]
moduleGenes=moduleGenes[-c(6,8,12,14)]

l=load("../Data/ReadyData/ReadyData-second-italy.RData")
dat_Expr_1 <- olink 
groups_1 <- samples_second[match(row.names(dat_Expr_1),samples_second$SampleID),"modifiedType",drop=TRUE]
l=load("../Data/ReadyData/ReadyData-third-heparin-nonTB.RData")
dat_Expr_2 <- olink 
groups_2 <- samples_third[match(row.names(dat_Expr_2),samples_third$SampleID),"modifiedType",drop=TRUE]

first_dataset <- dat_Expr_1%>%
  select(moduleGenes)%>%
  mutate(Class=groups_1)

first_dataset <- first_dataset%>%
  filter(Class%in%c("Active"))

second_dataset <- dat_Expr_2%>%
  select(moduleGenes)%>%
  mutate(Class="nonTB")

dataset <- first_dataset%>%
  rbind(second_dataset)

dataset$Class <- factor(dataset$Class,levels=c("Active","nonTB"))
dataset$Sample=as.factor(row.names(dataset))
data <- dataset
```

### Run random forest classification by 5-fold cross validation and compute ROC for each fold

```{r }
set.seed(123)
data <- fold(data, k = 5, cat_col = "Class", id_col = 'Sample')


rocs <- list()
confmat <- list()
counts <- list()
prs=list()
# One iteration per fold
for (fold in 1:5){
  
  # Create training set for this iteration
  # Subset all the datapoints where .folds does not match the current fold
  training_set <- data[data$.folds != fold,]
  
  # Create test set for this iteration
  # Subset all the datapoints where .folds matches the current fold
  testing_set <- data[data$.folds == fold,]
  training=subset(as.data.frame(training_set),select=-c(Sample,.folds))
  testing=subset(as.data.frame(testing_set),select=-c(Sample,.folds))
  
  set.seed(123)
  fitControl <- trainControl(method = "repeatedcv",
                             ## 10-fold CV...
                             repeats = 10,
                             number = 10,
                             classProbs = TRUE,
                             summaryFunction = defaultSummary)
  
   set.seed(123)
   model <- train(Class ~ ., data = training, 
                 method = "ranger", 
                 trControl = fitControl,metric="Kappa")
  
  
  pred <- predict(model, testing)
  prob <- predict(model, testing, type="prob")
  mat0 <- confusionMatrix(pred,testing$Class)
  
  roc0 <- roc(response = testing$Class, 
              predictor = prob[,1], 
              levels = rev(levels(testing$Class)),ci=TRUE)
  pr<-pr.curve(scores.class0 = testing$Class, weights.class0 = prob[,1],curve=TRUE);
  
  rocs[[paste0("fold:",fold)]]=roc0
  prs[[paste0("fold:",fold)]]=pr
  confmat[[paste0("fold:",fold)]]=mat0
  
  counts[[paste0("fold:",fold)]]=c(sum(testing$Class == "nonTB"),nrow(testing))
}

plot(prs[[1]])
```

### Aggregate ROC and PR plots of different folds into one main plot 

```{r}
get_proc <- function(roc_obj,counts_obj){
  list(
    "fpr" = 1 - roc_obj$specificities,
    "tpr" = roc_obj$sensitivities,
    "thresholds" = roc_obj$thresholds,
    "negative_count"= counts_obj[1],
    "total_count" = counts_obj[2],
    "auc" = roc_obj$auc
  )
}

roc_obj_node_1 <- get_proc(rocs[[1]],counts[[1]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_2 <- get_proc(rocs[[2]],counts[[2]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_3 <- get_proc(rocs[[3]],counts[[3]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_4 <- get_proc(rocs[[4]],counts[[4]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_5 <- get_proc(rocs[[5]],counts[[5]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases

# Preparing the input
fpr <- list(roc_obj_node_1$fpr, roc_obj_node_2$fpr, roc_obj_node_3$fpr,roc_obj_node_4$fpr,roc_obj_node_5$fpr)
tpr <- list(roc_obj_node_1$tpr, roc_obj_node_2$tpr, roc_obj_node_3$tpr,roc_obj_node_4$tpr,roc_obj_node_5$tpr)
thresholds <- list(
  roc_obj_node_1$thresholds, roc_obj_node_2$thresholds, roc_obj_node_3$thresholds,roc_obj_node_4$thresholds,roc_obj_node_5$thresholds)
negative_count <- c(
  roc_obj_node_1$negative_count, roc_obj_node_2$negative_count, roc_obj_node_3$negative_count,roc_obj_node_4$negative_count,roc_obj_node_5$negative_count)
total_count <- c(
  roc_obj_node_1$total_count, roc_obj_node_2$total_count, roc_obj_node_3$total_count,roc_obj_node_4$total_count,roc_obj_node_5$total_count)

# Compute the global ROC curve for the model
roc_aggregated <- roc_curve(fpr, tpr, thresholds, negative_count, total_count)

roc_aggregated$sensitivities <- roc_aggregated$tpr
roc_aggregated$specificities <- 1-roc_aggregated$fpr
# Calculate the AUC
roc_auc <- trapz(roc_aggregated$fpr, roc_aggregated$tpr)

sprintf("ROC AUC aggregated from each node's results: %f", roc_auc)

precision_recall_aggregated <- precision_recall_curve(
  fpr, tpr, thresholds, negative_count, total_count)
# Calculate the precision-recall AUC
precision_recall_auc <- -trapz(
  precision_recall_aggregated$recall, precision_recall_aggregated$pre)

sprintf(
  "Precision-Recall AUC aggregated from each node's results: %f",
  precision_recall_auc
)
```

### Visualize results using ROC plot

```{r}
path_cols <- brewer.pal(8, "Set1")[c(1,2,4,5)]

new_rocs=data.frame()
j=1
roc_folds <- function(roc_obj){
  new_rocs <<- rbind(new_rocs,data.frame(tpr=roc_obj$sensitivities,fpr=1-roc_obj$specificities,fold=paste0("Fold ",as.character(j)),color="gray"))
  new_rocs <<- new_rocs[order(new_rocs$fold,new_rocs$fpr,new_rocs$tpr,decreasing = FALSE),]
  j <<- j+1
}
lapply(rocs,function(x) roc_folds(x))
rocPlot<- data.frame(tpr=roc_aggregated$tpr,fpr=roc_aggregated$fpr,fold="Average of Folds",color="notGray")

Fig9_A <- ggplot(new_rocs,aes(x = fpr, y = tpr,group=fold))+
  geom_line(aes(linetype=fold),colour='gray')+
  geom_line(data=rocPlot, aes(x = fpr, y = tpr),linetype=1,linewidth=2)+
  labs(x = "1 - Specificity", y = "Sensitivity")+
  ggtitle(paste0("Average AUC=",format(round(roc_auc,2),nsmall=2)))+
  scale_linetype_manual(values=c(2,3,4,5,6))+
  theme_bw()+
  theme(plot.title = element_text(size=9),axis.title = element_text(size=12),axis.text = element_text(size=12))+
  guides(linetype=FALSE)+
  xlim(0,1)+
  ylim(0,1)
Fig9_A
```


## Signature including 6 proteins

### Load data for TB and nonTB from Italy

```{r message=FALSE, warning=FALSE}
l=load("../Data/ReadyData/AllDatasets-Samples.RData")
moduleGenes=c("IL6","CDCP1","IFN-gamma","CXCL9","VEGFA","MCP-3")
l=load("../Data/ReadyData/ReadyData-second-italy.RData")
dat_Expr_1 <- olink 
groups_1 <- samples_second[match(row.names(dat_Expr_1),samples_second$SampleID),"modifiedType",drop=TRUE]
l=load("../Data/ReadyData/ReadyData-third-heparin-nonTB.RData")
dat_Expr_2 <- olink 
groups_2 <- samples_third[match(row.names(dat_Expr_2),samples_third$SampleID),"modifiedType",drop=TRUE]

first_dataset <- dat_Expr_1%>%
  select(moduleGenes)%>%
  mutate(Class=groups_1)

first_dataset <- first_dataset%>%
  filter(Class%in%c("Active"))

second_dataset <- dat_Expr_2%>%
  select(moduleGenes)%>%
  mutate(Class="nonTB")

dataset <- first_dataset%>%
  rbind(second_dataset)

dataset$Class <- factor(dataset$Class,levels=c("Active","nonTB"))
dataset$Sample=as.factor(row.names(dataset))
data <- dataset
```

### Run random forest classification by 5-fold cross validation and compute ROC for each fold

```{r }
set.seed(123)
data <- fold(data, k = 5, cat_col = "Class", id_col = 'Sample')


rocs <- list()
confmat <- list()
counts <- list()
prs=list()
# One iteration per fold
for (fold in 1:5){
  
  # Create training set for this iteration
  # Subset all the datapoints where .folds does not match the current fold
  training_set <- data[data$.folds != fold,]
  
  # Create test set for this iteration
  # Subset all the datapoints where .folds matches the current fold
  testing_set <- data[data$.folds == fold,]
  training=subset(as.data.frame(training_set),select=-c(Sample,.folds))
  testing=subset(as.data.frame(testing_set),select=-c(Sample,.folds))
  
  set.seed(123)
  fitControl <- trainControl(method = "repeatedcv",
                             ## 10-fold CV...
                             repeats = 10,
                             number = 10,
                             classProbs = TRUE,
                             summaryFunction = defaultSummary)
  
   set.seed(123)
   model <- train(Class ~ ., data = training, 
                 method = "ranger", 
                 trControl = fitControl,metric="Kappa")
  
  
  pred <- predict(model, testing)
  prob <- predict(model, testing, type="prob")
  mat0 <- confusionMatrix(pred,testing$Class)
  
  roc0 <- roc(response = testing$Class, 
              predictor = prob[,1], 
              levels = rev(levels(testing$Class)),ci=TRUE)
  pr<-pr.curve(scores.class0 = testing$Class, weights.class0 = prob[,1],curve=TRUE);
  
  rocs[[paste0("fold:",fold)]]=roc0
  prs[[paste0("fold:",fold)]]=pr
  confmat[[paste0("fold:",fold)]]=mat0
  
  counts[[paste0("fold:",fold)]]=c(sum(testing$Class == "nonTB"),nrow(testing))
}
```

### Aggregate ROC and PR plots of different folds into one main plot 

```{r}
get_proc <- function(roc_obj,counts_obj){
  list(
    "fpr" = 1 - roc_obj$specificities,
    "tpr" = roc_obj$sensitivities,
    "thresholds" = roc_obj$thresholds,
    "negative_count"= counts_obj[1],
    "total_count" = counts_obj[2],
    "auc" = roc_obj$auc
  )
}

roc_obj_node_1 <- get_proc(rocs[[1]],counts[[1]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_2 <- get_proc(rocs[[2]],counts[[2]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_3 <- get_proc(rocs[[3]],counts[[3]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_4 <- get_proc(rocs[[4]],counts[[4]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_5 <- get_proc(rocs[[5]],counts[[5]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases

# Preparing the input
fpr <- list(roc_obj_node_1$fpr, roc_obj_node_2$fpr, roc_obj_node_3$fpr,roc_obj_node_4$fpr,roc_obj_node_5$fpr)
tpr <- list(roc_obj_node_1$tpr, roc_obj_node_2$tpr, roc_obj_node_3$tpr,roc_obj_node_4$tpr,roc_obj_node_5$tpr)
thresholds <- list(
  roc_obj_node_1$thresholds, roc_obj_node_2$thresholds, roc_obj_node_3$thresholds,roc_obj_node_4$thresholds,roc_obj_node_5$thresholds)
negative_count <- c(
  roc_obj_node_1$negative_count, roc_obj_node_2$negative_count, roc_obj_node_3$negative_count,roc_obj_node_4$negative_count,roc_obj_node_5$negative_count)
total_count <- c(
  roc_obj_node_1$total_count, roc_obj_node_2$total_count, roc_obj_node_3$total_count,roc_obj_node_4$total_count,roc_obj_node_5$total_count)

# Compute the global ROC curve for the model
roc_aggregated <- roc_curve(fpr, tpr, thresholds, negative_count, total_count)

roc_aggregated$sensitivities <- roc_aggregated$tpr
roc_aggregated$specificities <- 1-roc_aggregated$fpr
# Calculate the AUC
roc_auc <- trapz(roc_aggregated$fpr, roc_aggregated$tpr)

sprintf("ROC AUC aggregated from each node's results: %f", roc_auc)

precision_recall_aggregated <- precision_recall_curve(
  fpr, tpr, thresholds, negative_count, total_count)
# Calculate the precision-recall AUC
precision_recall_auc <- -trapz(
  precision_recall_aggregated$recall, precision_recall_aggregated$pre)

sprintf(
  "Precision-Recall AUC aggregated from each node's results: %f",
  precision_recall_auc
)
```

### Visualize results using ROC plot

```{r}
new_rocs=data.frame()
j=1
roc_folds <- function(roc_obj){
  new_rocs <<- rbind(new_rocs,data.frame(tpr=roc_obj$sensitivities,fpr=1-roc_obj$specificities,fold=paste0("Fold ",as.character(j)),color="gray"))
  new_rocs <<- new_rocs[order(new_rocs$fold,new_rocs$fpr,new_rocs$tpr,decreasing = FALSE),]
  j <<- j+1
}
lapply(rocs,function(x) roc_folds(x))
rocPlot<- data.frame(tpr=roc_aggregated$tpr,fpr=roc_aggregated$fpr,fold="Average of Folds",color="notGray")


Fig9_B <- ggplot(new_rocs,aes(x = fpr, y = tpr,group=fold))+
  geom_line(aes(linetype=fold),colour='gray')+
  geom_line(data=rocPlot, aes(x = fpr, y = tpr),linetype=1,linewidth=2)+
  labs(x = "1 - Specificity", y = "Sensitivity")+
  ggtitle(paste0("Average AUC=",format(round(roc_auc,2),nsmall=2)))+
  scale_linetype_manual(values=c(2,3,4,5,6))+
  theme_bw()+
  theme(plot.title = element_text(size=9),axis.title = element_text(size=12),axis.text = element_text(size=12))+
  guides(linetype=FALSE)+
  xlim(0,1)+
  ylim(0,1)
Fig9_B
```



## Signature including 4 proteins

### Load data for TB and nonTB from Italy

```{r message=FALSE, warning=FALSE}
l=load("../Data/ReadyData/AllDatasets-Samples.RData")
moduleGenes=c("CDCP1","IFN-gamma","CXCL9","VEGFA")
l=load("../Data/ReadyData/ReadyData-second-italy.RData")
dat_Expr_1 <- olink 
groups_1 <- samples_second[match(row.names(dat_Expr_1),samples_second$SampleID),"modifiedType",drop=TRUE]
l=load("../Data/ReadyData/ReadyData-third-heparin-nonTB.RData")
dat_Expr_2 <- olink 
groups_2 <- samples_third[match(row.names(dat_Expr_2),samples_third$SampleID),"modifiedType",drop=TRUE]

first_dataset <- dat_Expr_1%>%
  select(moduleGenes)%>%
  mutate(Class=groups_1)

first_dataset <- first_dataset%>%
  filter(Class%in%c("Active"))

second_dataset <- dat_Expr_2%>%
  select(moduleGenes)%>%
  mutate(Class="nonTB")

dataset <- first_dataset%>%
  rbind(second_dataset)

dataset$Class <- factor(dataset$Class,levels=c("Active","nonTB"))
dataset$Sample=as.factor(row.names(dataset))
data <- dataset
```

### Run random forest classification by 5-fold cross validation and compute ROC for each fold

```{r }
set.seed(123)
data <- fold(data, k = 5, cat_col = "Class", id_col = 'Sample')


rocs <- list()
confmat <- list()
counts <- list()
prs=list()
# One iteration per fold
for (fold in 1:5){
  
  # Create training set for this iteration
  # Subset all the datapoints where .folds does not match the current fold
  training_set <- data[data$.folds != fold,]
  
  # Create test set for this iteration
  # Subset all the datapoints where .folds matches the current fold
  testing_set <- data[data$.folds == fold,]
  training=subset(as.data.frame(training_set),select=-c(Sample,.folds))
  testing=subset(as.data.frame(testing_set),select=-c(Sample,.folds))
  
  set.seed(123)
  fitControl <- trainControl(method = "repeatedcv",
                             ## 10-fold CV...
                             repeats = 10,
                             number = 10,
                             classProbs = TRUE,
                             summaryFunction = defaultSummary)
  
   set.seed(123)
   model <- train(Class ~ ., data = training, 
                 method = "ranger", 
                 trControl = fitControl,metric="Kappa")
  
  
  pred <- predict(model, testing)
  prob <- predict(model, testing, type="prob")
  mat0 <- confusionMatrix(pred,testing$Class)
  
  roc0 <- roc(response = testing$Class, 
              predictor = prob[,1], 
              levels = rev(levels(testing$Class)),ci=TRUE)
  pr<-pr.curve(scores.class0 = testing$Class, weights.class0 = prob[,1],curve=TRUE);
  
  rocs[[paste0("fold:",fold)]]=roc0
  prs[[paste0("fold:",fold)]]=pr
  confmat[[paste0("fold:",fold)]]=mat0
  
  counts[[paste0("fold:",fold)]]=c(sum(testing$Class == "nonTB"),nrow(testing))
}
```

### Aggregate ROC and PR plots of different folds into one main plot 

```{r}
get_proc <- function(roc_obj,counts_obj){
  list(
    "fpr" = 1 - roc_obj$specificities,
    "tpr" = roc_obj$sensitivities,
    "thresholds" = roc_obj$thresholds,
    "negative_count"= counts_obj[1],
    "total_count" = counts_obj[2],
    "auc" = roc_obj$auc
  )
}

roc_obj_node_1 <- get_proc(rocs[[1]],counts[[1]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_2 <- get_proc(rocs[[2]],counts[[2]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_3 <- get_proc(rocs[[3]],counts[[3]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_4 <- get_proc(rocs[[4]],counts[[4]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases
roc_obj_node_5 <- get_proc(rocs[[5]],counts[[5]])
#> Setting levels: control = 0, case = 1
#> Setting direction: controls < cases

# Preparing the input
fpr <- list(roc_obj_node_1$fpr, roc_obj_node_2$fpr, roc_obj_node_3$fpr,roc_obj_node_4$fpr,roc_obj_node_5$fpr)
tpr <- list(roc_obj_node_1$tpr, roc_obj_node_2$tpr, roc_obj_node_3$tpr,roc_obj_node_4$tpr,roc_obj_node_5$tpr)
thresholds <- list(
  roc_obj_node_1$thresholds, roc_obj_node_2$thresholds, roc_obj_node_3$thresholds,roc_obj_node_4$thresholds,roc_obj_node_5$thresholds)
negative_count <- c(
  roc_obj_node_1$negative_count, roc_obj_node_2$negative_count, roc_obj_node_3$negative_count,roc_obj_node_4$negative_count,roc_obj_node_5$negative_count)
total_count <- c(
  roc_obj_node_1$total_count, roc_obj_node_2$total_count, roc_obj_node_3$total_count,roc_obj_node_4$total_count,roc_obj_node_5$total_count)

# Compute the global ROC curve for the model
roc_aggregated <- roc_curve(fpr, tpr, thresholds, negative_count, total_count)

roc_aggregated$sensitivities <- roc_aggregated$tpr
roc_aggregated$specificities <- 1-roc_aggregated$fpr
# Calculate the AUC
roc_auc <- trapz(roc_aggregated$fpr, roc_aggregated$tpr)

sprintf("ROC AUC aggregated from each node's results: %f", roc_auc)

precision_recall_aggregated <- precision_recall_curve(
  fpr, tpr, thresholds, negative_count, total_count)
# Calculate the precision-recall AUC
precision_recall_auc <- -trapz(
  precision_recall_aggregated$recall, precision_recall_aggregated$pre)

sprintf(
  "Precision-Recall AUC aggregated from each node's results: %f",
  precision_recall_auc
)
```

### Visualize results using ROC plot

```{r}
new_rocs=data.frame()
j=1
roc_folds <- function(roc_obj){
  new_rocs <<- rbind(new_rocs,data.frame(tpr=roc_obj$sensitivities,fpr=1-roc_obj$specificities,fold=paste0("Fold ",as.character(j)),color="gray"))
  new_rocs <<- new_rocs[order(new_rocs$fold,new_rocs$fpr,new_rocs$tpr,decreasing = FALSE),]
  j <<- j+1
}
lapply(rocs,function(x) roc_folds(x))
rocPlot<- data.frame(tpr=roc_aggregated$tpr,fpr=roc_aggregated$fpr,fold="Average of Folds",color="notGray")

Fig9_C <- ggplot(new_rocs,aes(x = fpr, y = tpr,group=fold))+
  geom_line(aes(linetype=fold),colour='gray')+
  geom_line(data=rocPlot, aes(x = fpr, y = tpr),linetype=1,linewidth=2)+
  labs(x = "1 - Specificity", y = "Sensitivity")+
  ggtitle(paste0("Average AUC=",format(round(roc_auc,2),nsmall=2)))+
  scale_linetype_manual(values=c(2,3,4,5,6))+
  theme_bw()+
  theme(plot.title = element_text(size=9),axis.title = element_text(size=12),axis.text = element_text(size=12))+
  guides(linetype=FALSE)+
  xlim(0, 1)+
  ylim(0,1)
Fig9_C
```

# Save results

```{r}
save(Fig9_A,Fig9_B,Fig9_C,file="../Results/Figures/Fig9.RData")
```

