---
title: "New Olink Study"
author: "Zaynab Mousavian"
date: "7 August 2024"
output:
  html_document: default
  pdf_document: default
subtitle: Making feagures
---

```{r setup, include=FALSE}
# Set global options for all chunks
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Loading R packages

```{r}
library(ggplot2)
library(patchwork)
library(ggpubr)
library(grid)
library(ComplexHeatmap)
```

# Function for removing redundant information including yaxis labels and legend 

```{r}
remove_axis_titles <- function(plot) {
  plot +
    theme(
      axis.title.y = element_blank()
    )
}
remove_axis_titles_x <- function(plot) {
  plot +
    theme(
      axis.title.x = element_blank()
    )
}
remove_legend <- function(plot) {
  plot +
    theme(
      legend.position = "None"
    )
}
```

# Load all figures

```{r}
folder_path <- "../Results/Figures/"
rdata_files <- list.files(path = folder_path, pattern = "\\.RData$", full.names = TRUE)

# Load each .RData file
for (file in rdata_files) {
  load(file)
}
```

# Making figure 2 for 12-marker

```{r}
Fig1_ssGSEA_discovery <- remove_legend(Fig1_ssGSEA_discovery)
Fig1_ssGSEA_both <- remove_legend(Fig1_ssGSEA_both)
Fig1_ssGSEA_sweden <- remove_legend(Fig1_ssGSEA_sweden)
Fig1_ssGSEA_both <- remove_axis_titles(Fig1_ssGSEA_both)
Fig1_ssGSEA_sweden <- remove_axis_titles(Fig1_ssGSEA_sweden)
Fig1_ssGSEA_italy <- remove_axis_titles(Fig1_ssGSEA_italy)
Fig1_ssGSEA <- Fig1_ssGSEA_discovery+ggtitle("A: Discovery cohort (Training)")+Fig1_ssGSEA_both+ggtitle("B: Validation cohort (Testing)")+Fig1_ssGSEA_sweden+ggtitle("C: Subcohort_1 (Sweden)")+Fig1_ssGSEA_italy+ggtitle("D: Subcohort_2 (Italy)")+
  plot_layout(ncol=4)
Fig1_ROCtext_discovery <- remove_legend(Fig1_ROCtext_discovery)
Fig1_ROCtext_both <- remove_legend(Fig1_ROCtext_both)
Fig1_ROCtext_sweden <- remove_legend(Fig1_ROCtext_sweden)
Fig1_ROCtext_both <- remove_axis_titles(Fig1_ROCtext_both)
Fig1_ROCtext_sweden <- remove_axis_titles(Fig1_ROCtext_sweden)
Fig1_ROCtext_italy <- remove_axis_titles(Fig1_ROCtext_italy)
Fig1_ROC <- Fig1_ROCtext_discovery+Fig1_ROCtext_both+Fig1_ROCtext_sweden+Fig1_ROCtext_italy+
  plot_layout(ncol=4)
Fig1 <- (Fig1_ssGSEA/Fig1_ROC)+
  plot_layout(heights = c(1, 1)) 
ggsave(Fig1,file="../Results/Figures/Manuscript/Figure 1.pdf",width=15, height = 10)
```

# Making supp figure 1 showing subtypes of active TB

```{r}
suppFig1_ssGSEA_sweden_subTypes_A <- remove_legend(suppFig1_ssGSEA_sweden_subTypes_A)
suppFig1_ssGSEA_sweden_subTypes_B <- remove_axis_titles(suppFig1_ssGSEA_sweden_subTypes_B)
suppFig1 <- suppFig1_ssGSEA_sweden_subTypes_A + suppFig1_ssGSEA_sweden_subTypes_B+
  plot_layout(ncol=2)
ggsave(suppFig1,file="../Results/Figures/Manuscript/Figure S1.pdf",width=10,height = 5)
```

# Making supp figure 2 for correlation with clinical variables

```{r}
load("../Results/Figures/Fig7_correlationplot.RData")
legend_theme <- theme(legend.position = "bottom", 
                      legend.direction = "horizontal")

correlationplot_CRP_A <- remove_legend(correlationplot_CRP_A)
correlationplot_CRP_A <- correlationplot_CRP_A+ggtitle("12-marker")
correlationplot_ESR_A <- remove_legend(correlationplot_ESR_A)
correlationplot_HB_A <- remove_legend(correlationplot_HB_A)
correlationplot_ALB_A <- remove_legend(correlationplot_ALB_A)

Fig7 <- correlationplot_CRP_A+correlationplot_ESR_A+correlationplot_HB_A+correlationplot_ALB_A+correlationplot_WBC_A+
  plot_layout(widths = c(1,1,1,1,1))

Fig7 <- Fig7+legend_theme
ggsave(Fig7,file="../Results/Figures/Manuscript/Figure S2.pdf", width=20,height = 7)
```


# Making Figure 3 for 6 markers and best 4 markers 

```{r}
Fig2_ssGSEA_discovery_B <- remove_legend(Fig2_ssGSEA_discovery_B)
Fig2_ssGSEA_both_B <- remove_legend(Fig2_ssGSEA_both_B)
Fig2_ssGSEA_sweden_B <- remove_legend(Fig2_ssGSEA_sweden_B)
Fig2_ssGSEA_both_B <- remove_axis_titles(Fig2_ssGSEA_both_B)
Fig2_ssGSEA_sweden_B <- remove_axis_titles(Fig2_ssGSEA_sweden_B)
Fig2_ssGSEA_italy_B <- remove_axis_titles(Fig2_ssGSEA_italy_B)
Fig2_ssGSEA_B <- Fig2_ssGSEA_discovery_B+ggtitle("A: Discovery cohort (Training)")+Fig2_ssGSEA_both_B+ggtitle("B: Validation cohort (Testing)")+Fig2_ssGSEA_sweden_B+ggtitle("C: Subcohort_1 (Sweden)")+Fig2_ssGSEA_italy_B+ggtitle("D: Subcohort_2 (Italy)")+
  plot_layout(ncol=4)

Fig2_ROCtext_discovery_B <- remove_legend(Fig2_ROCtext_discovery_B)
Fig2_ROCtext_both_B <- remove_legend(Fig2_ROCtext_both_B)
Fig2_ROCtext_sweden_B <- remove_legend(Fig2_ROCtext_sweden_B)
Fig2_ROCtext_both_B <- remove_axis_titles(Fig2_ROCtext_both_B)
Fig2_ROCtext_sweden_B <- remove_axis_titles(Fig2_ROCtext_sweden_B)
Fig2_ROCtext_italy_B <- remove_axis_titles(Fig2_ROCtext_italy_B)
Fig2_ROCtext_B <- Fig2_ROCtext_discovery_B+Fig2_ROCtext_both_B+Fig2_ROCtext_sweden_B+Fig2_ROCtext_italy_B+
  plot_layout(ncol=4)

Fig2_ssGSEA_discovery_F <- remove_legend(Fig2_ssGSEA_discovery_F)
Fig2_ssGSEA_both_F <- remove_legend(Fig2_ssGSEA_both_F)
Fig2_ssGSEA_sweden_F <- remove_legend(Fig2_ssGSEA_sweden_F)
Fig2_ssGSEA_both_F <- remove_axis_titles(Fig2_ssGSEA_both_F)
Fig2_ssGSEA_sweden_F <- remove_axis_titles(Fig2_ssGSEA_sweden_F)
Fig2_ssGSEA_italy_F <- remove_axis_titles(Fig2_ssGSEA_italy_F)
Fig2_ssGSEA_F <- Fig2_ssGSEA_discovery_F+ggtitle("E: Discovery cohort (Training)")+Fig2_ssGSEA_both_F+ggtitle("F: Validation cohort (Testing)")+Fig2_ssGSEA_sweden_F+ggtitle("G: Subcohort_1 (Sweden)")+Fig2_ssGSEA_italy_F+ggtitle("H: Subcohort_2 (Italy)")+
  plot_layout(ncol=4)

Fig2_ROCtext_discovery_F <- remove_legend(Fig2_ROCtext_discovery_F)
Fig2_ROCtext_both_F <- remove_legend(Fig2_ROCtext_both_F)
Fig2_ROCtext_sweden_F <- remove_legend(Fig2_ROCtext_sweden_F)
Fig2_ROCtext_both_F <- remove_axis_titles(Fig2_ROCtext_both_F)
Fig2_ROCtext_sweden_F <- remove_axis_titles(Fig2_ROCtext_sweden_F)
Fig2_ROCtext_italy_F <- remove_axis_titles(Fig2_ROCtext_italy_F)
Fig2_ROCtext_F <- Fig2_ROCtext_discovery_F+Fig2_ROCtext_both_F+Fig2_ROCtext_sweden_F+Fig2_ROCtext_italy_F+
  plot_layout(ncol=4)

Fig2 <- (Fig2_ssGSEA_B/Fig2_ROCtext_B/Fig2_ssGSEA_F/Fig2_ROCtext_F)+
  plot_layout(heights = c(1, 1,1,1)) 

ggsave(Fig2,file="../Results/Figures/Manuscript/Figure 2.pdf",width=15, height = 20)
```

# Making supp figure 4 for 7 markers from Boruta

```{r}
Fig2_ssGSEA_discovery_A <- remove_legend(Fig2_ssGSEA_discovery_A)
Fig2_ssGSEA_both_A <- remove_legend(Fig2_ssGSEA_both_A)
Fig2_ssGSEA_sweden_A <- remove_legend(Fig2_ssGSEA_sweden_A)
Fig2_ssGSEA_both_A <- remove_axis_titles(Fig2_ssGSEA_both_A)
Fig2_ssGSEA_sweden_A <- remove_axis_titles(Fig2_ssGSEA_sweden_A)
Fig2_ssGSEA_italy_A <- remove_axis_titles(Fig2_ssGSEA_italy_A)
Fig2_ssGSEA_A <- Fig2_ssGSEA_discovery_A+ggtitle("A: Discovery cohort (Training)")+Fig2_ssGSEA_both_A+ggtitle("B: Validation cohort (Testing)")+Fig2_ssGSEA_sweden_A+ggtitle("C: Subcohort_1 (Sweden)")+Fig2_ssGSEA_italy_A+ggtitle("D: Subcohort_2 (Italy)")+
  plot_layout(ncol=4)

Fig2_ROCtext_discovery_A <- remove_legend(Fig2_ROCtext_discovery_A)
Fig2_ROCtext_both_A <- remove_legend(Fig2_ROCtext_both_A)
Fig2_ROCtext_sweden_A <- remove_legend(Fig2_ROCtext_sweden_A)
Fig2_ROCtext_both_A <- remove_axis_titles(Fig2_ROCtext_both_A)
Fig2_ROCtext_sweden_A <- remove_axis_titles(Fig2_ROCtext_sweden_A)
Fig2_ROCtext_italy_A <- remove_axis_titles(Fig2_ROCtext_italy_A)
Fig2_ROCtext_A <- Fig2_ROCtext_discovery_A+Fig2_ROCtext_both_A+Fig2_ROCtext_sweden_A+Fig2_ROCtext_italy_A+
  plot_layout(ncol=4)

suppFig3 <- (Fig2_ssGSEA_A/Fig2_ROCtext_A)+
  plot_layout(heights = c(1, 1)) 

ggsave(suppFig3,file="../Results/Figures/Manuscript/Figure S4.pdf",width=15, height = 10)
```

# Making supp figure 6 for all combinations of 4 out of 6

```{r}
combined_plot_discovery_4marker <- remove_legend(combined_plot_discovery_4marker)
combined_plot_both_4marker<- remove_legend(combined_plot_both_4marker)
combined_plot_sweden_4marker<- remove_legend(combined_plot_sweden_4marker)
combined_plot_both_4marker <- remove_axis_titles(combined_plot_both_4marker)
combined_plot_sweden_4marker <- remove_axis_titles(combined_plot_sweden_4marker)
combined_plot_italy_4marker <- remove_axis_titles(combined_plot_italy_4marker)
suppFig5 <- (combined_plot_discovery_4marker+ggtitle("A: Discovery cohort (Training)")+combined_plot_both_4marker+ggtitle("B: Validation cohort (Testing)")+combined_plot_sweden_4marker+ggtitle("C: Subcohort_1 (Sweden)")+combined_plot_italy_4marker)+ggtitle("D: Subcohort_2 (Italy)")+
  plot_layout(ncol=4)
ggsave(suppFig5,file="../Results/Figures/Manuscript/Figure S6.pdf",width=20, height = 5)
```


# Making supp figure 7 

```{r}
Fig5_ssGSEA_sweden <- remove_legend(Fig5_ssGSEA_sweden)
Fig5_ssGSEA_sweden_B<- remove_legend(Fig5_ssGSEA_sweden_B)
Fig5_ssGSEA_sweden_B <- remove_axis_titles(Fig5_ssGSEA_sweden_B)
Fig5_ssGSEA_sweden_F <- remove_axis_titles(Fig5_ssGSEA_sweden_F)

Fig5_ssGSEA_sweden <- Fig5_ssGSEA_sweden+Fig5_ssGSEA_sweden_B+Fig5_ssGSEA_sweden_F+
  plot_layout(ncol=3)

Fig5_correlationplot <- remove_legend(Fig5_correlationplot)
Fig5_correlationplot_B<- remove_legend(Fig5_correlationplot_B)
Fig5_correlationplot_B <- remove_axis_titles(Fig5_correlationplot_B)
Fig5_correlationplot_F <- remove_axis_titles(Fig5_correlationplot_F)
Fig5_correlation <- Fig5_correlationplot+ggtitle("")+Fig5_correlationplot_B+ggtitle("")+Fig5_correlationplot_F+ggtitle("")+
  plot_layout(ncol=3)
Fig5 <- (Fig5_ssGSEA_sweden/Fig5_correlation)+
  plot_layout(heights = c(1,1))
ggsave(Fig5,file="../Results/Figures/Manuscript/Figure S7.pdf",width = 10,height = 10)
```

# Making supp figure 8 for correlation between EDTA and heparin in individual marker

```{r}
suppFig11 <- Fig5_correlationplot_markers
ggsave(suppFig11,file="../Results/Figures/Manuscript/Figure S8.pdf",width = 15,height = 15)
```


# Making supp figure 9 for all individual markers

```{r}
ggsave(Fig5_scatterplot_markers, file="../Results/Figures/Manuscript/Figure S9.pdf",width = 10,height = 10)
```

# Making figure 4

```{r}
Fig3_ssGSEA <- remove_legend(Fig3_ssGSEA)
Fig3_ssGSEA_B<- remove_legend(Fig3_ssGSEA_B)
Fig3_ssGSEA_B <- remove_axis_titles(Fig3_ssGSEA_B)
Fig3_ssGSEA_F <- remove_axis_titles(Fig3_ssGSEA_F)
Fig3_GSEA <- Fig3_ssGSEA+Fig3_ssGSEA_B+Fig3_ssGSEA_F

heatmap_grob <- grid.grabExpr(draw(Fig3_heatmap))

Fig3 <- Fig3_GSEA/ wrap_elements(heatmap_grob)+ggtitle("D")
ggsave(Fig3,file="../Results/Figures/Manuscript/Figure 3.pdf",width = 10,height = 10)
```


# Making supp figure 10 for non TB ROC (cross validation on whole non TB and TB italy)

```{r}
Fig9_B <- remove_axis_titles(Fig9_B)
Fig9_C <- remove_axis_titles(Fig9_C)
Fig9 <- Fig9_A+Fig9_B+Fig9_C
ggsave(Fig9,file="../Results/Figures/Manuscript/Figure S10.pdf", width=10,height = 3)
```

# Making figure 5 for comparison with other signatures

```{r}
combined_plot_discovery_comparison <- remove_legend(combined_plot_discovery_comparison)
combined_plot_both_comparison<- remove_legend(combined_plot_both_comparison)
combined_plot_sweden_comparison<- remove_legend(combined_plot_sweden_comparison)
combined_plot_comparison <-combined_plot_discovery_comparison+ggtitle("A: Discovery cohort (Training)")+combined_plot_both_comparison+ggtitle("B: Validation cohort (Testing)")+combined_plot_sweden_comparison+ggtitle("C: Subcohort_1 (Sweden)")+combined_plot_italy_comparison+ggtitle("D: Subcohort_2 (Italy)")+
  plot_layout(ncol=4) 
Fig6_B <- remove_legend(Fig6_ssGSEA_TB+ggtitle("E"))+Fig6_pValue_TB+ggtitle("F")+
  plot_layout(widths = c(5,0.5))
Fig6_C <- remove_legend(Fig6_ssGSEA_nonTB+ggtitle("G"))+Fig6_pValue_nonTB+ggtitle("H")+
  plot_layout(widths = c(5,0.5))
Fig6 <- combined_plot_comparison/Fig6_B/Fig6_C+
  plot_layout(heights = c(1,1,1),widths = c(1,1,1))
ggsave(Fig6,file="../Results/Figures/Manuscript/Figure 4.pdf", width=15,height = 15)
```

# Making supp figure 11 for all individual markers

```{r}
suppFig2 <- (((suppFig2_ROCtext_discovery_IL6+ggtitle("A: Discovery cohort (Training)") + suppFig2_ROCtext_both_IL6+ggtitle("B: Validation cohort (Testing)")+ suppFig2_ROCtext_sweden_IL6+ggtitle("C: Subcohort_1 (Sweden)")+suppFig2_ROCtext_italy_IL6+ggtitle("D: Subcohort_2 (Italy)")+suppFig2_ROCtext_discovery_CDCP1+ggtitle("E: Discovery cohort (Training)") + suppFig2_ROCtext_both_CDCP1+ggtitle("F: Validation cohort (Testing)")+ suppFig2_ROCtext_sweden_CDCP1+ggtitle("G: Subcohort_1 (Sweden)")+suppFig2_ROCtext_italy_CDCP1+ggtitle("H: Subcohort_2 (Italy)"))+plot_layout(ncol=8))/
  ((suppFig2_ROCtext_discovery_VEGFA + suppFig2_ROCtext_both_VEGFA+suppFig2_ROCtext_sweden_VEGFA+suppFig2_ROCtext_italy_VEGFA+suppFig2_ROCtext_discovery_MCP3 +suppFig2_ROCtext_both_MCP3+ suppFig2_ROCtext_sweden_MCP3+suppFig2_ROCtext_italy_MCP3)+plot_layout(ncol=8))/
  ((suppFig2_ROCtext_discovery_IFNG + suppFig2_ROCtext_both_IFNG+suppFig2_ROCtext_sweden_IFNG+suppFig2_ROCtext_italy_IFNG+suppFig2_ROCtext_discovery_CXCL9 + suppFig2_ROCtext_both_CXCL9+ suppFig2_ROCtext_sweden_CXCL9+suppFig2_ROCtext_italy_CXCL9)+plot_layout(ncol=8))/
  ((suppFig2_ROCtext_discovery_CCL19 + suppFig2_ROCtext_both_CCL19+suppFig2_ROCtext_sweden_CCL19+suppFig2_ROCtext_italy_CCL19+suppFig2_ROCtext_discovery_MMP1 + suppFig2_ROCtext_both_MMP1+ suppFig2_ROCtext_sweden_MMP1+suppFig2_ROCtext_italy_MMP1)+plot_layout(ncol=8))/
((suppFig2_ROCtext_discovery_IL7 + suppFig2_ROCtext_both_IL7+suppFig2_ROCtext_sweden_IL7+suppFig2_ROCtext_italy_IL7+suppFig2_ROCtext_discovery_IL12B +suppFig2_ROCtext_both_IL12B+ suppFig2_ROCtext_sweden_IL12B+suppFig2_ROCtext_italy_IL12B)+plot_layout(ncol=8))/
  ((suppFig2_ROCtext_discovery_CD40 + suppFig2_ROCtext_both_CD40+suppFig2_ROCtext_sweden_CD40+suppFig2_ROCtext_italy_CD40+suppFig2_ROCtext_discovery_PDL1 +suppFig2_ROCtext_both_PDL1 + suppFig2_ROCtext_sweden_PDL1+suppFig2_ROCtext_italy_PDL1)+plot_layout(ncol=8)))+
  plot_layout(nrow=6)

ggsave(suppFig2,file="../Results/Figures/Manuscript/Figure S11.pdf",width=20, height = 25)
```

