---
title: "New Olink Study"
author: "Zaynab Mousavian"
date: "12 June 2024"
output:
  html_document: default
  pdf_document: default
subtitle: Bridge samples
---

```{r setup, include=FALSE}
# Set global options for all chunks
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Loading R packages

```{r}
library(OlinkAnalyze)
library(readxl)
library(tidyverse)
```

# 1. Read all datasets

```{r}
npx_data_June19 <- read_NPX(filename = "../Data/Olink iscience/June-2019/NPX_data_files/Stimulated TB PBMCs (plasma samples)_NPX_LOD.xlsx")
npx_data_Nov19 <- read_NPX(filename = "../Data/Olink iscience/Nov-2019/NPX_data_files/MTB_plasma_profiling_NPX_LOD.xlsx")
npx_data_Jan20 <- read_NPX("../Data/Olink iscience/Jan-2020/NPX_data_files/TB_5_days_supernatant_NPX_LOD.xlsx")
npx_data_May22 <- read_NPX(filename = "../Data/Olink old validation/Plasma profiling May 2022/shared_with_user/NPX_data_files/TB validation_NPX_LOD.xlsx")
npx_data_edta <- read_NPX(filename ="../Data/edta_Data_NPX_LOD_wide.xlsx")
npx_data_heparin <- read_NPX(filename ="../Data/heparin_Data_NPX_LOD_wide.xlsx")
load("../Data/ReadyData/AllDatasets-Samples.RData")
```

# 2. Bridge three datasets in the first dataset

```{r}
npx_data_June19 <- read_NPX(filename = "../Data/Olink iscience/June-2019/NPX_data_files/Stimulated TB PBMCs (plasma samples)_NPX_LOD.xlsx")
npx_data_Nov19 <- read_NPX(filename = "../Data/Olink iscience/Nov-2019/NPX_data_files/MTB_plasma_profiling_NPX_LOD.xlsx")
npx_data_Jan20 <- read_NPX("../Data/Olink iscience/Jan-2020/NPX_data_files/TB_5_days_supernatant_NPX_LOD.xlsx")

# remove samples with NA values
npx_data_Nov19 <- npx_data_Nov19 %>% 
  filter(!SampleID%in%c("LS040", "LS042")) 

samples <- data.frame(read_excel("../Data/Olink iscience/Samples-discovery.xlsx"))

npx_data_June19 <- npx_data_June19 %>% left_join(samples, by = "SampleID")
npx_data_Nov19 <- npx_data_Nov19 %>% left_join(samples, by = "SampleID")
npx_data_Jan20 <- npx_data_Jan20 %>% left_join(samples, by = "SampleID")

npx_data_Jan20 <- npx_data_Jan20%>%
  filter(!is.na(subjectID))

npx_data_June19$SampleID <- npx_data_June19$subjectID
npx_data_Nov19$SampleID <- npx_data_Nov19$subjectID
npx_data_Jan20$SampleID <- npx_data_Jan20$subjectID

npx_data1 <- npx_data_June19
npx_data2 <- npx_data_Nov19
npx_data3 <- npx_data_Jan20

npx_1 <- npx_data1 %>%
  mutate(Project = "data1")
npx_2 <- npx_data2 %>%
  mutate(Project = "data2")
npx_3 <- npx_data3 %>%
  mutate(Project = "data3")

overlap_samples <-data.frame(SampleID = intersect(npx_1$SampleID, npx_2$SampleID)) %>%
  filter(!str_detect(SampleID, "CONTROL_SAMPLE")) %>% #Remove control samples
  pull(SampleID)

overlap_samples_list <- list("DF1" = overlap_samples,
                             "DF2" = overlap_samples)

# Perform Bridging normalization
npx_br_data <- olink_normalization_bridge(project_1_df = npx_1,
                                          project_2_df = npx_2,
                                          bridge_samples = overlap_samples_list,
                                          project_1_name = "data1",
                                          project_2_name = "data2",
                                          project_ref_name = "data1")

new_normalized_data_1 <- npx_br_data %>% 
  dplyr::filter(Project == "data1") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_2 <- npx_br_data %>% 
  dplyr::filter(Project == "data2") %>% 
  dplyr::select(-Project, -Adj_factor) 

overlap_samples <-data.frame(SampleID = intersect(npx_1$SampleID, npx_3$SampleID)) %>%
  filter(!str_detect(SampleID, "CONTROL_SAMPLE")) %>% #Remove control samples
  pull(SampleID)

overlap_samples_list <- list("DF1" = overlap_samples,
                             "DF2" = overlap_samples)

# Perform Bridging normalization
npx_br_data <- olink_normalization_bridge(project_1_df = npx_1,
                                          project_2_df = npx_3,
                                          bridge_samples = overlap_samples_list,
                                          project_1_name = "data1",
                                          project_2_name = "data3",
                                          project_ref_name = "data1")

new_normalized_data_1 <- npx_br_data %>% 
  dplyr::filter(Project == "data1") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_3 <- npx_br_data %>% 
  dplyr::filter(Project == "data3") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_1 <- new_normalized_data_1%>%
  select(-Batch,-Group,-Origin,-subjectID)
new_normalized_data_2 <- new_normalized_data_2%>%
  select(-Batch,-Group,-Origin,-subjectID)
new_normalized_data_3 <- new_normalized_data_3%>%
  select(-Batch,-Group,-Origin,-subjectID)
```

# 3. Bridge second dataset with first dataset 

```{r}
npx_data_May22 <- read_NPX(filename = "../Data/Olink old validation/Plasma profiling May 2022/shared_with_user/NPX_data_files/TB validation_NPX_LOD.xlsx")

npx_data_May22 <- npx_data_May22 %>% 
  filter(!SampleID%in%c("Facility Ctrl 1", "Facility Ctrl 2","LAM007 (2)","LAM008 (2)","DS61","DS13", "TB0061","TB0059")) %>%
  select(-Normalization)


npx_data1 <- new_normalized_data_1
npx_data2 <- npx_data_May22

npx_1 <- npx_data1 %>%
  mutate(Project = "data1")
npx_2 <- npx_data2 %>%
  mutate(Project = "data2")

overlap_samples_list <-list("DF1" = c("LS003", "LS004","AS007_R","LS008","AS009_R","HS012","HS013","HS016"),
                           "DF2" = c("LAM003","LAM004","LAM007 (1)","LAM008 (1)","LAM009","LAM012","LAM013","LAM016"))

npx_br_data <- olink_normalization_bridge(project_1_df = npx_1,
                                          project_2_df = npx_2,
                                          bridge_samples = overlap_samples_list,
                                          project_1_name = "data1",
                                          project_2_name = "data2",
                                          project_ref_name = "data1")

new_normalized_data_1 <- npx_br_data %>% 
  dplyr::filter(Project == "data1") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_4 <- npx_br_data %>% 
  dplyr::filter(Project == "data2") %>% 
  dplyr::select(-Project, -Adj_factor) 
```


# 4. Bridge dataset3-heparin with second dataset

```{r}
npx_data_heparin <- read_NPX(filename ="../Data/heparin_Data_NPX_LOD_wide.xlsx")

npx_data_heparin <- npx_data_heparin %>%
  filter(!SampleID%in%c("Facility Ctrl", "Facility Ctrl-2", "MaxLOD","PZ 3")) %>%
  select(-Normalization)

npx_data_edta <- read_NPX(filename ="../Data/edta_Data_NPX_LOD_wide.xlsx")

npx_data_edta <- npx_data_edta %>% 
  filter(!SampleID%in%c("Facility Ctrl", "Facility Ctrl-2", "MaxLOD","TB0133.e","TB0161.e")) %>%
  select(-Normalization)

npx_data1 <- new_normalized_data_4
npx_data2 <- npx_data_heparin%>%rbind(npx_data_edta)

npx_1 <- npx_data1 %>%
  mutate(Project = "data1")
npx_2 <- npx_data2 %>%
  mutate(Project = "data2")

overlap_samples_list <-list("DF1" = c("STAMIR","DIFRA","RAKIMI","XIBIN","SERIU","LUKORA","CIFAB3","FORTAN","TB0075","TB0077","TB0078","TB0079","TB0089","TB0091","TB0092","TB0096","TB0102","TB0108","TB0109"),
                           "DF2" = c("Old delia1", "Old delia2", "Old delia3", "Old delia4","Old delia5", "Old delia6", "Old delia7", "Old delia8","TB0075.0","TB0077.0","TB0078","TB0079","TB0089","TB0091.0","TB0092.0","TB0096","TB0102","TB0108","TB0109"))

npx_br_data <- olink_normalization_bridge(project_1_df = npx_1,
                                          project_2_df = npx_2,
                                          bridge_samples = overlap_samples_list,
                                          project_1_name = "data1",
                                          project_2_name = "data2",
                                          project_ref_name = "data1")


new_normalized_data_4 <- npx_br_data %>% 
  dplyr::filter(Project == "data1") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_5 <- npx_br_data %>% 
  dplyr::filter(Project == "data2") %>% 
  dplyr::select(-Project, -Adj_factor) 
```

# 5. Bridge dataset3-heparin with first dataset 

```{r}
npx_data1 <- new_normalized_data_2
npx_data2 <- new_normalized_data_5

npx_1 <- npx_data1 %>%
  mutate(Project = "data1")
npx_2 <- npx_data2 %>%
  mutate(Project = "data2")

overlap_samples_list <-list("DF1" = c("AS035", "AS029"),
                           "DF2" = c("TB0035.0","TB0029.0"))

npx_br_data <- olink_normalization_bridge(project_1_df = npx_1,
                                          project_2_df = npx_2,
                                          bridge_samples = overlap_samples_list,
                                          project_1_name = "data1",
                                          project_2_name = "data2",
                                          project_ref_name = "data1")


new_normalized_data_2 <- npx_br_data %>% 
  dplyr::filter(Project == "data1") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_5 <- npx_br_data %>% 
  dplyr::filter(Project == "data2") %>% 
  dplyr::select(-Project, -Adj_factor) 

npx_data1 <- new_normalized_data_3
npx_data2 <- new_normalized_data_5

npx_1 <- npx_data1 %>%
  mutate(Project = "data1")
npx_2 <- npx_data2 %>%
  mutate(Project = "data2")

overlap_samples_list <-list("DF1" = c("AS0038","AS1002"),
                           "DF2" = c("TB0038.0","TBHA1002.0"))

npx_br_data <- olink_normalization_bridge(project_1_df = npx_1,
                                          project_2_df = npx_2,
                                          bridge_samples = overlap_samples_list,
                                          project_1_name = "data1",
                                          project_2_name = "data2",
                                          project_ref_name = "data1")

new_normalized_data_3 <- npx_br_data %>% 
  dplyr::filter(Project == "data1") %>% 
  dplyr::select(-Project, -Adj_factor) 

new_normalized_data_5 <- npx_br_data %>% 
  dplyr::filter(Project == "data2") %>% 
  dplyr::select(-Project, -Adj_factor) 
```

# 6. Showing all datasets using PCA plot {.tabset}

## Before bridge

```{r message=FALSE}
npx_data_June19 <- npx_data_June19 %>% select(-Batch,-Group,-Origin,-subjectID)
npx_data_Nov19 <- npx_data_Nov19 %>% select(-Batch,-Group,-Origin,-subjectID)
npx_data_Jan20 <- npx_data_Jan20 %>% select(-Batch,-Group,-Origin,-subjectID)

npx_before_br <- npx_data_June19%>%
  mutate(SampleID=paste0(SampleID,"-data1"), Type="data1")%>%
  rbind((npx_data_Nov19%>%
 mutate(SampleID=paste0(SampleID,"-data2"), Type="data2")))%>%
  rbind((npx_data_Jan20%>%
  mutate(SampleID=paste0(SampleID,"-data3"), Type="data3")))%>%
  rbind((npx_data_May22%>%
  mutate(SampleID=paste0(SampleID,"-data4"), Type="data4")))%>%
  rbind((npx_data_heparin%>%
 mutate(SampleID=paste0(SampleID,"-data5"), Type="data5")))

### PCA plot
OlinkAnalyze::olink_pca_plot(df  = npx_before_br,
                             color_g     = "Type",
                             byPanel     = TRUE)
```

## After bridge

```{r message=FALSE}
npx_after_br <- new_normalized_data_1%>%
  mutate(SampleID=paste0(SampleID,"-data1"), Type="data1")%>%
  rbind((new_normalized_data_2%>%
 mutate(SampleID=paste0(SampleID,"-data2"), Type="data2")))%>%
  rbind((new_normalized_data_3%>%
  mutate(SampleID=paste0(SampleID,"-data3"), Type="data3")))%>%
  rbind((new_normalized_data_4%>%
  mutate(SampleID=paste0(SampleID,"-data4"), Type="data4")))%>%
  rbind((new_normalized_data_5%>%
 mutate(SampleID=paste0(SampleID,"-data5"), Type="data5")))

### PCA plot
OlinkAnalyze::olink_pca_plot(df  = npx_after_br,
                             color_g     = "Type",
                             byPanel     = TRUE,
                             label_outliers = TRUE)
```

# Showing groups to see the effect of bridging {.tabset}

## Before bridge

```{r}
samples <- read_excel(path="../Data/Plasma profiling_Feb_2024_final.xlsx",sheet = "Samples and randomization")
samples$`TB status` <- gsub("Latent","Latent remote",samples$`TB status`)
samples$`TB status` <- gsub("Latent remote remote","Latent remote",samples$`TB status`)
samples$`TB status` <- gsub("Remote LTBI","Latent remote",samples$`TB status`)
samples <- samples %>% rename("SampleID"=`Sample ID`,"Group"=`TB status`) 

npx_data_heparin_2 <- npx_data_heparin %>% left_join(samples,by = "SampleID")
npx_data_edta_2 <- npx_data_edta %>% left_join(samples,by = "SampleID")

samples <- data.frame(read_excel("../Data/Olink old validation/Samples-validation-old run.xlsx"))
samples <- samples %>% rename("SampleID"=Sample.ID,"Group"=Group) 

npx_data_May22_2 <- npx_data_May22 %>% left_join(samples,by = "SampleID")

npx_data_heparin_2  <- npx_data_heparin_2 %>% select(-`Study ID`,-Rack,-Box,-...6,-...7,-...8)
npx_data_edta_2  <- npx_data_edta_2 %>% select(-`Study ID`,-Rack,-Box,-...6,-...7,-...8)
npx_data_May22_2  <- npx_data_May22_2 %>% select(-Sample.origin,-Diagnosis,-Final.group,-...5)

npx_before_br <- npx_data_heparin_2 %>%
  mutate(SampleID=paste0(SampleID,"-data5"), Type="data5")%>%
  arrange(Group)%>%
  rbind((npx_data_edta_2 %>%
  mutate(SampleID=paste0(SampleID,"-data5"), Type="data5")%>%
  arrange(Group)))

OlinkAnalyze::olink_pca_plot(df  = npx_before_br,
                             color_g = "Group",
                             byPanel     = TRUE)

npx_before_br <- npx_data_May22_2 %>%
  mutate(SampleID=paste0(SampleID,"-data4"), Type="data4")%>%
  arrange(Group)

OlinkAnalyze::olink_pca_plot(df  = npx_before_br,
                             color_g = "Group",
                             byPanel     = TRUE)

npx_before_br <- npx_data_heparin_2 %>%
  mutate(SampleID=paste0(SampleID,"-data5"), Type="data5")%>%
  arrange(Group)%>%
  rbind((npx_data_edta_2 %>%
  mutate(SampleID=paste0(SampleID,"-data5"), Type="data5")%>%
  arrange(Group)))%>%
  rbind((npx_data_May22_2 %>%
  mutate(SampleID=paste0(SampleID,"-data4"), Type="data4")%>%
  arrange(Group)))

OlinkAnalyze::olink_pca_plot(df  = npx_before_br,
                             color_g = "Group",
                             byPanel     = TRUE)
```

## After bridge

```{r}
samples <- read_excel(path="../Data/Plasma profiling_Feb_2024_final.xlsx",sheet = "Samples and randomization")
samples$`TB status` <- gsub("Latent","Latent remote",samples$`TB status`)
samples$`TB status` <- gsub("Latent remote remote","Latent remote",samples$`TB status`)
samples$`TB status` <- gsub("Remote LTBI","Latent remote",samples$`TB status`)
samples <- samples %>% rename("SampleID"=`Sample ID`,"Group"=`TB status`) 

new_normalized_data_5_2 <- new_normalized_data_5 %>%
  left_join(samples,by = "SampleID")

samples <- data.frame(read_excel("../Data/Olink old validation/Samples-validation-old run.xlsx"))
samples <- samples %>% rename("SampleID"=Sample.ID,"Group"=Group) 

new_normalized_data_4_2 <- new_normalized_data_4 %>%
  left_join(samples,by = "SampleID")

new_normalized_data_5_2  <- new_normalized_data_5_2 %>% select(-`Study ID`,-Rack,-Box,-...6,-...7,-...8)
new_normalized_data_4_2  <- new_normalized_data_4_2 %>% select(-Sample.origin,-Diagnosis,-Final.group,-...5)

npx_after_br <- new_normalized_data_5_2 %>%
  mutate(SampleID=paste0(SampleID,"-data5"), Type="data5") %>%
  arrange(Group)

OlinkAnalyze::olink_pca_plot(df  = npx_after_br,
                              color_g = "Group",
                             byPanel     = TRUE)

npx_after_br <- new_normalized_data_4_2 %>%
  mutate(SampleID=paste0(SampleID,"-data4"), Type="data4")%>%
  arrange(Group)
### PCA plot
OlinkAnalyze::olink_pca_plot(df  = npx_after_br,
                              color_g = "Group",
                             byPanel     = TRUE)

npx_after_br <- new_normalized_data_5_2 %>%
  mutate(SampleID=paste0(SampleID,"-data5"), Type="data5") %>%
  arrange(Group)%>%
  rbind((new_normalized_data_4_2 %>%
  mutate(SampleID=paste0(SampleID,"-data4"), Type="data4")%>%
  arrange(Group)))

OlinkAnalyze::olink_pca_plot(df  = npx_after_br,
                              color_g = "Group",
                             byPanel     = TRUE)
```

# Make data ready {.tabset}

## First dataset

```{r}
npx_data_first <- new_normalized_data_1%>%
  rbind((new_normalized_data_2))%>%
  rbind(new_normalized_data_3)

samples_sweden <- samples_first%>%
  filter(Origin=="Sweden")%>%
  select(SampleID)%>%
  unlist()
npx_data_first_sweden <- npx_data_first%>%
  filter(SampleID%in%samples_sweden)

samples_portugal <- samples_first%>%
  filter(Origin=="Portugal")%>%
  select(SampleID)%>%
  unlist()
npx_data_first_portugal <- npx_data_first%>%
  filter(SampleID%in%samples_portugal)

samples_first <- samples_first%>%
  mutate(modifiedType = case_when(
    str_detect(Type,"Healthy controls") ~ "Control",
    str_detect(Type,"Contacts") ~ "Latent",
    TRUE ~ Type))

samples_first <- samples_first%>%
  filter(!is.na(modifiedType))

samples_first%>%
  select(modifiedType)%>%
  table()
```

## Second dataset

```{r}
npx_data_second <- new_normalized_data_4
overlapping_samples <- c("LAM003","LAM004","LAM007 (1)","LAM008 (1)","LAM009","LAM012","LAM013","LAM016")

samples_sweden <- samples_second%>%
  filter(`Sample origin`=="Sweden")%>%
  select(SampleID)%>%
  unlist()
npx_data_second_sweden <- npx_data_second%>%
  filter(SampleID%in%samples_sweden)%>%
  filter(!SampleID%in%overlapping_samples)

samples_italy <- samples_second%>%
  filter(`Sample origin`=="Italy")%>%
  select(SampleID)%>%
  unlist()
npx_data_second_italy <- npx_data_second%>%
  filter(SampleID%in%samples_italy)%>%
  filter(!SampleID%in%overlapping_samples)

samples_second <- samples_second%>%
  mutate(modifiedType = case_when(
    str_detect(Type,"Active") ~ "Active",
    str_detect(Type,"Aktiv") ~ "Active",
    str_detect(Type,"aktiv") ~ "Active",
    str_detect(Type,"PULMONARY TB") ~ "Active",
    str_detect(Type,"active") ~ "Active",
    str_detect(Type,"Latent") ~ "Latent",
    str_detect(Type,"latent") ~ "Latent",
    str_detect(Type,"LTBI") ~ "Latent",
    str_detect(Type,"IGRA+ contact") ~ "Latent",
    str_detect(Type,"Contact, IGRA+") ~ "Latent",
    str_detect(Type,"Contact, IGRA-") ~ "Control",
    str_detect(Type,"IGRA- contact") ~ "Control",
    str_detect(Type,"Control, IGRA-") ~ "Control",
    str_detect(Type,"HD") ~ "Control",
    str_detect(`Final group`,"Latent") ~ "Latent",
    str_detect(`Final group`,"Control") ~ "Control",
    TRUE ~ Type))

samples_second[which(samples_second$SampleID=="TB0068"),"modifiedType"]="Control"
samples_second[which(samples_second$SampleID=="TB0088"),"modifiedType"]="Control"
samples_second[which(samples_second$SampleID=="TB0091"),"modifiedType"]="Active"

samples_second <- samples_second%>%
  filter(!is.na(modifiedType))

samples_second%>%
  select(modifiedType)%>%
  table()
```

## Third dataset

```{r}
heparin_samples <- npx_data_heparin%>%select(SampleID)%>%unique()%>%unlist()
edta_samples <- npx_data_edta%>%select(SampleID)%>%unique()%>%unlist()
overlapping_samples <- c("Old delia1", "Old delia2", "Old delia3", "Old delia4","Old delia5", "Old delia6", "Old delia7", "Old delia8","TB0075.0","TB0077.0","TB0078","TB0079","TB0089","TB0091.0","TB0092.0","TB0096","TB0102","TB0108","TB0109","TB0035.0","TB0029.0","TB0038.0","TBHA1002.0")

new_normalized_data_5_heparin <- new_normalized_data_5%>%
  filter(SampleID%in%heparin_samples)%>%
  filter(!SampleID%in%overlapping_samples)

new_normalized_data_5_edta <- new_normalized_data_5%>%
  filter(SampleID%in%edta_samples)

new_normalized_data_5_heparin <- new_normalized_data_5_heparin %>%
  mutate(ends_with_x = str_detect(SampleID, "_r$")) %>%
  filter(ends_with_x==FALSE)

new_normalized_data_5_edta <- new_normalized_data_5_edta %>%
  mutate(ends_with_x = str_detect(SampleID, "_r$")) %>%
  filter(ends_with_x==FALSE)

npx_data_third_heparin <- new_normalized_data_5_heparin

npx_data_third_heparin_nonTB <- npx_data_third_heparin%>%
  filter(str_starts(SampleID, "PZ"))

npx_data_third_heparin_followTreatment <- npx_data_third_heparin%>%
  filter(str_ends(SampleID,"\\.0"))%>%
  rbind((npx_data_third_heparin%>%
           filter(str_ends(SampleID, "\\.1"))))%>%
  rbind((npx_data_third_heparin%>%
           filter(str_ends(SampleID, "\\.2"))))

npx_data_third_heparin <- npx_data_third_heparin%>%
  filter(!str_starts(SampleID,"PZ"),!str_ends(SampleID, "\\.1"),!str_ends(SampleID, "\\.2"))
  
npx_data_third_edta <- new_normalized_data_5_edta

samples_third <- samples_third%>%
  mutate(modifiedType = case_when(
    str_detect(Type,"Active") ~ "Active",
    str_detect(Type,"Pulmonary TB") ~ "Active",
    str_detect(Type,"Latent") ~ "Latent",
    str_detect(Type,"LTBI") ~ "Latent",
    str_detect(Type,"HC") ~ "Control",
    TRUE~ Type))

samples_third <- samples_third%>%
  mutate(Treatment = case_when(
    str_ends(SampleID,"\\.0") ~ "Baseline",
    str_ends(SampleID,"\\.1") ~ "2 Months",
    str_ends(SampleID,"\\.2") ~ "6 Months",
    TRUE ~ "Non-treatment"))


samples_third%>%
  select(modifiedType)%>%
  table()

samples_third%>%
  select(Treatment)%>%
  table()
```

# Remove eyeTBs

```{r}
npx_data_third_heparin <- npx_data_third_heparin%>%
  filter(SampleID!="TB0198")

npx_data_third_heparin_followTreatment <- npx_data_third_heparin_followTreatment%>%
  filter(SampleID!="TB0077.1"&SampleID!="TB0077.2")

npx_data_third_edta <- npx_data_third_edta%>%
  filter(SampleID!="TB0077.e")

npx_data_third_edta <- npx_data_third_edta%>%
  filter(SampleID!="TB0198.e")
```

# Save all datasets

```{r}
save(npx_data_first_sweden,npx_data_first_portugal,npx_data_second_sweden,npx_data_second_italy,npx_data_third_heparin,npx_data_third_heparin_nonTB,npx_data_third_heparin_followTreatment,npx_data_third_edta, file="../Data/ReadyData/AllDatasets-NPX.RData")

write.table(samples_first,file="../Results/Samples_first.txt",sep = "\t",row.names = FALSE)
write.table(samples_second,file="../Results/Samples_second.txt",sep = "\t",row.names = FALSE)
write.table(samples_third,file="../Results/Samples_third.txt",sep = "\t",row.names = FALSE)

save(samples_first, samples_second, samples_third,file="../Data/ReadyData/AllDatasets-Samples.RData")
```

