---
title: "New Olink Study"
author: "Zaynab Mousavian"
date: "3 June 2024"
output:
  html_document: default
  pdf_document: default
subtitle: Olink preprocessing
---

```{r setup, include=FALSE}
# Set global options for all chunks
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Loading R packages

```{r}
library(OlinkAnalyze)
library(readxl)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
```

# Read samples information

```{r}
l=load("../Data/ReadyData/AllDatasets-NPX.RData")
l=load("../Data/ReadyData/AllDatasets-Samples.RData")
```

# Which protein data? {.tabset}

## discovery

### Read protein data 

```{r}
npx_data_first_sweden%>%
  glimpse()
  
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_first_sweden %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```
### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_third_edta%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_first[match(row.names(filteredOlink),samples_first$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_first$modifiedType))
colors <- brewer.pal(9, "Set3")

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_first[match(row.names(filteredOlink),samples_first$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_first$modifiedType))

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-discovery.RData")
```




## Portugal

### Read protein data 

```{r}
npx_data_first_portugal%>%
  glimpse()
  
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_first_portugal %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```
### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_third_edta%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_first[match(row.names(filteredOlink),samples_first$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_first$modifiedType))
colors <- brewer.pal(9, "Set3")
# If you need more than the maximu
# pdf(file=paste0(resultsDirectory,"Figure S1-edta.pdf"))
ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))
# dev.off()

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_first[match(row.names(filteredOlink),samples_first$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_first$modifiedType))
#pdf(file=paste0(resultsDirectory,"Figure S2-edta.pdf"))
ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))
#dev.off()

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-Portugal.RData")
```




## Edta

### Read protein data 

```{r}
npx_data_third_edta%>%
  glimpse()
  
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_third_edta %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```
### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_third_edta%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$modifiedType))
colors <- brewer.pal(9, "Set3")
# If you need more than the maximu
# pdf(file=paste0(resultsDirectory,"Figure S1-edta.pdf"))
ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))
# dev.off()

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$modifiedType))

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-third-edta.RData")
```




## Heparin

### Read protein data 

```{r}
npx_data_third_heparin%>%
  glimpse()
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_third_heparin %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```
### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_third_heparin%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$modifiedType))
colors <- brewer.pal(9, "Set3")

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$modifiedType))
#pdf(file=paste0(resultsDirectory,"Figure S2-edta.pdf"))
ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725", "#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))
#dev.off()

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-third-heparin.RData")
```




## Heparin TB treatment

### Read protein data 

```{r}
npx_data_third_heparin_followTreatment%>%
  glimpse()
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_third_heparin_followTreatment %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```
### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_third_heparin_followTreatment%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"Treatment"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","Treatment"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$Treatment))
colors <- brewer.pal(9, "Set3")

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = colors)+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))
# dev.off()

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"Treatment"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","Treatment"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$Treatment))

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = colors)+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-third-heparin-followTreatment.RData")
```




## Heparin nonTB

### Read protein data 

```{r}
npx_data_third_heparin_nonTB%>%
  glimpse()
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_third_heparin_nonTB %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```
### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_third_heparin_nonTB%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$modifiedType))
colors <- brewer.pal(9, "Set1")

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = colors)+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_third[match(row.names(filteredOlink),samples_third$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = unique(samples_third$modifiedType))

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = colors)+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-third-heparin-nonTB.RData")
```




## Second sweden

### Read protein data 

```{r}
npx_data_second_sweden%>%
  glimpse()
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_second_sweden %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```

### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_second_sweden%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_second[match(row.names(filteredOlink),samples_second$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = c("Active","Latent","Control"))
colors <- brewer.pal(9, "Set1")

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725","#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_second[match(row.names(filteredOlink),samples_second$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = c("Active","Latent","Control"))

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725","#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-second-sweden.RData")
```




## Second italy

### Read protein data 

```{r}
npx_data_second_italy%>%
  glimpse()
```

### Convert protein data from long to wide format

```{r}
olink <- npx_data_second_italy %>%
  select(SampleID,Assay,NPX)
olink <- dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), 
               fun.aggregate=function(i) mean(i, na.rm=TRUE))

rownames(olink) <- olink$SampleID
olink <- olink%>%
  select(-SampleID)
```

### Filtering proteins 

```{r message=FALSE}
percentage <- 0.3
cutoff <- nrow(olink)*percentage
nSamples <- nrow(olink)
# Get LOD; since LOD varies between panels, collect individual LOD per protein then average them out
LOD <- npx_data_second_italy%>%
  select(Assay,Panel_Version,LOD)
unique(LOD$Panel_Version)
LOD <- as.data.frame(LOD[!duplicated(LOD$Assay),])
LOD <- LOD%>%select(-Panel_Version)
LOD$Assay <- as.character(LOD$Assay)
rownames(LOD) <- LOD$Assay
identical(LOD$Assay,colnames(olink)) #TRUE
# Add LOD to olink df
olink <- as.data.frame(t(olink))
olink <- merge(olink, LOD[,2, drop=FALSE], by="row.names", all = TRUE)
rownames(olink) <- olink$Row.names
olink$Row.names <- NULL
colnames(olink)[nSamples+1] <- "LOD"
# Detect number of samples with limit of detection (LOD)
num_LOD <- as.data.frame(apply(olink, 1, function(x) length(which(x[1:nSamples] <= x[nSamples+1])) ))
olink <- merge(olink, num_LOD, by="row.names")
rownames(olink) <- olink$Row.names
olink <- olink[,-1]
colnames(olink)[nSamples+2] <- "n_LOD"
olink <- olink[,-(nSamples+1)] #remove LOD
olink <- as.data.frame(t(olink))

# Remove proteins from analysis where more than 30% of samples have NPX values less than LOD 
filteredProteins <- which(olink[nSamples+1,] > cutoff)
filteredOlink <- olink[-(nSamples+1),filteredProteins]
filteredOlink <- cbind(filteredOlink,samples_second[match(row.names(filteredOlink),samples_second$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = c("Active","Latent","Control"))
colors <- brewer.pal(9, "Set1")

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725","#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=7)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

# Remaining proteins with less than 30% of samples have NPX values less than LOD 
remainingProteins <- which(olink[nSamples+1,] <= cutoff)
filteredOlink <- olink[-(nSamples+1),remainingProteins]
filteredOlink <- cbind(filteredOlink,samples_second[match(row.names(filteredOlink),samples_second$SampleID),"modifiedType"])
filteredOlink$sample <- row.names(filteredOlink)
inputData <- melt(filteredOlink,id=c("sample","modifiedType"))
names(inputData) <- c("Samples", "Groups", "Proteins", "NPX")
inputData$Groups <- factor(inputData$Groups, levels = c("Active","Latent","Control"))

ggplot(inputData,aes(x=Proteins,y=NPX,fill=Groups))+
  geom_boxplot(aes(middle=mean(NPX)),fatten=1,outlier.size = 0.5)+
  scale_fill_manual(values = c("#2F8AC4", "#E48725","#A5AA99"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
  facet_wrap(~Proteins, scale="free",ncol=8)+
  theme(axis.text=element_text(size=6),
        axis.text.x=element_blank(),
        axis.title = element_text(size=7,face="bold"),
        strip.text = element_text(size=7),
        legend.title = element_blank(), 
        legend.text = element_text(size=7))

olink <- olink[,-filteredProteins] # 
olink <- olink[-(nSamples+1),] #remove lod_n info
```


### Save protein data

```{r}
save(olink,file="../Data/ReadyData/ReadyData-second-italy.RData")
```


