---
title: "New Olink Study"
author: "Zaynab Mousavian"
date: "8 August 2024"
output:
  html_document: default
  pdf_document: default
subtitle: Correlation plots
---

```{r setup, include=FALSE}
# Set global options for all chunks
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Loading R packages

```{r}
library(OlinkAnalyze)
library(readxl)
library(dplyr)
library(reshape2)
library(ggplot2)
library(corto)
library(patchwork)
library(ggpubr)
library(tidyverse)
```

# Scatterplot for correlation with facet

```{r}
scatterplot_func_withfacet <- function(df,x,y,group,colors){
  correlations <- df %>%
  group_by(.data[[group]]) %>%
  summarize(correlation = cor(.data[[x]], .data[[y]], method = "pearson"))

  plot <- ggplot(data=df, aes(x = .data[[x]], y = .data[[y]], color=colors)) +
    geom_point() +  
    geom_smooth(method = "lm", se = FALSE, color = "gray",size=0.5) +  
    scale_color_manual(values = c("#2F8AC4", "#E48725"),labels = c("Active" = "TB Disease","Latent" = "TB Infection"))+
    facet_wrap(~ .data[[group]],scales = "free") +  
    labs(title = "", x = "EDTA", y = "Heparin",color="")+
    theme_minimal()+
    theme(axis.text.x = element_text(size = 9,angle=0),  
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size=12),
    panel.background = element_rect(fill = "white", color = NA),  
    plot.background = element_rect(fill = "white", color = NA),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )+
  geom_text(data = correlations, aes(x = Inf, y = Inf, label = paste0("r = ", round(correlation, 2))), 
            hjust = 1.1, vjust = 1.1, inherit.aes = FALSE, size = 3,color="darkgrey")
  
  return(plot)
}
```

# Scatterplot for correlation

```{r}
scatterplot_func <- function(df,x,y,type,x_axis,y_axis){
  
  overall_correlation <- cor(df[[x]], df[[y]], method = "pearson")
  
  plot <- ggplot(data=df, aes(x = .data[[x]], y = .data[[y]], color=.data[[type]])) +
    geom_point() +  
    geom_smooth(method = "lm", se = FALSE, color = "gray",size=0.5) +  
    scale_color_manual(values = c("#2F8AC4", "#E48725","gray"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
    labs(title = "", x = x_axis, y = y_axis,color="")+
    theme_minimal()+
    theme(axis.text.x = element_text(size = 9),  
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size=12),
    panel.background = element_rect(fill = "white", color = NA),  
    plot.background = element_rect(fill = "white", color = NA),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )+
    geom_text(aes(label = paste("Corr:", round(overall_correlation, 2))),
              x = -Inf, y = Inf, hjust = 0, vjust = 2, 
              size = 4, color = "darkgrey", inherit.aes = FALSE)
  
  return(plot)
}
```


# Scatterplot for correlation whithin each group

```{r}
scatterplot_func2 <- function(df,x,y,type,x_axis,y_axis){
  
  plot <- ggplot(data=df, aes(x = .data[[x]], y = .data[[y]], color=.data[[type]])) +
    geom_point() +  
    geom_smooth(method = "lm", se = FALSE, color = "gray",size=0.5) +  
    scale_color_manual(values = c("#2F8AC4", "#E48725","gray"),labels = c("Active" = "TB Disease","Latent" = "TB Infection","Control"="IGRA- Contact"))+
    labs(title = "", x = x_axis, y = y_axis,color="")+
    theme_minimal()+
    theme(axis.text.x = element_text(size = 9),  
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size=12),
    panel.background = element_rect(fill = "white", color = NA),  
    plot.background = element_rect(fill = "white", color = NA),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )+
  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = 'left', label.y.npc = 'top', 
           size = 3, 
           data = df[df[[type]] == "Active", ],color="#2F8AC4") + 

  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = 'left', label.y.npc = 'bottom', 
           size = 3, 
           data = df[df[[type]] == "Latent", ] ,color="#E48725")
  
  return(plot)
}
```

# load data
```{r}
l=load("../Data/ReadyData/AllDatasets-Samples.RData")
```

# load signature

```{r}
signature <- c("IFN-gamma","CXCL9","VEGFA","CDCP1","MCP-3","IL6","CCL19","CD40","IL7","MMP-1","PD-L1","IL-12B")
```

# read EDTA and Heparin NPX data

```{r}
load("../Data/ReadyData-withoutFilteringProteins/ReadyData-third-edta.RData")
dat_Expr_edta <- t(olink)
groups_edta <- samples_third[match(colnames(dat_Expr_edta),samples_third$SampleID),"modifiedType",drop=TRUE]

load("../Data/ReadyData-withoutFilteringProteins/ReadyData-third-heparin.RData")
dat_Expr_third_heparin <- t(olink)
load("../Data/ReadyData-withoutFilteringProteins/ReadyData-second-sweden.RData")
dat_Expr_second_sweden <- t(olink)

groups_third_heparin <- samples_third[match(colnames(dat_Expr_third_heparin),samples_third$SampleID),"modifiedType",drop=TRUE]
groups_second_sweden <- samples_second[match(colnames(dat_Expr_second_sweden),samples_second$SampleID),"modifiedType",drop=TRUE]

dat_Expr_heparin <- dat_Expr_third_heparin%>%
  cbind(dat_Expr_second_sweden)
groups_heparin <- c(groups_third_heparin,groups_second_sweden)
```

# Overlap between EDTA and Heparin data

```{r}
EDTA_samples <- colnames(dat_Expr_edta)
EDTA_samples <- gsub(".e","",EDTA_samples)
colnames(dat_Expr_edta) <- EDTA_samples
heparin_samples <- colnames(dat_Expr_heparin)
heparin_samples <- gsub("\\.0","",heparin_samples)
colnames(dat_Expr_heparin) <- heparin_samples

overlapping_samples <- intersect(heparin_samples,EDTA_samples)
dat_Expr_edta <- dat_Expr_edta[signature,overlapping_samples]
dat_Expr_heparin <- dat_Expr_heparin[signature,overlapping_samples]

samples_third$SampleID <- gsub(".e","",samples_third$SampleID )
groups <- samples_third[match(overlapping_samples,samples_third$SampleID),"modifiedType",drop=TRUE]
groups[41] <- "Latent"
```

# Visulizing correlation between EDTA and Heparin NPX values for the signature

```{r}
dat_Expr_edta_melted <- melt(dat_Expr_edta)
colnames(dat_Expr_edta_melted) <- c("Protein_EDTA","Sample_EDTA","NPX_EDTA")
dat_Expr_heparin_melted <- melt(dat_Expr_heparin)
colnames(dat_Expr_heparin_melted) <- c("Protein_Heparin","Sample_Heparin","NPX_Heparin")

input_data <- dat_Expr_edta_melted%>%
  cbind(dat_Expr_heparin_melted)

new_groups <- rep(groups,each=12)
new_groups <- factor(new_groups, levels = c("Active","Latent"))
  
Fig5_correlationplot_markers <- scatterplot_func_withfacet(input_data,"NPX_EDTA","NPX_Heparin","Protein_EDTA",new_groups)
Fig5_correlationplot_markers 
```

# Visulizing paired scatterplot between EDTA and Heparin NPX values for the signature

```{r}
dat_Expr_edta_melted <- melt(dat_Expr_edta)
colnames(dat_Expr_edta_melted) <- c("Protein","Sample","NPX")
dat_Expr_edta_melted$Anticoagulant <- rep("EDTA",nrow(dat_Expr_edta_melted))
dat_Expr_heparin_melted <- melt(dat_Expr_heparin)
colnames(dat_Expr_heparin_melted) <- c("Protein","Sample","NPX")
dat_Expr_heparin_melted$Anticoagulant <- rep("Heparin",nrow(dat_Expr_heparin_melted))

input_data <- dat_Expr_edta_melted%>%
  rbind(dat_Expr_heparin_melted)

new_groups <- rep(groups,each=12)
new_groups <- factor(new_groups, levels = c("Active","Latent"))
input_data$Type <- c(new_groups,new_groups)
input_data$Anticoagulant <- factor(input_data$Anticoagulant, levels = c("Heparin","EDTA"))

p_values <- input_data %>%
  group_by(Protein) %>%
  summarise(
    p_value = t.test(NPX[Anticoagulant == "Heparin"], 
                     NPX[Anticoagulant == "EDTA"], 
                     paired = TRUE)$p.value,
    mean_difference = t.test(NPX[Anticoagulant == "Heparin"], 
                     NPX[Anticoagulant == "EDTA"], 
                     paired = TRUE)$estimate,
  ) %>%
  mutate(
    p_label = ifelse(p_value < 0.001, "***",
                     ifelse(p_value < 0.01, "**",
                            ifelse(p_value < 0.05, "*", "ns"))),
    group1 = "Heparin",  # Specify first group
    group2 = "EDTA"      # Specify second group
  )

write.csv(p_values,file="../Results/Heparin-EDTA-markers.csv")
# Create plot
Fig5_scatterplot_markers <- ggplot(input_data, aes(x = Anticoagulant, y = NPX, color = Type, group = Sample)) +
  geom_point(size = 3) +                         
  geom_line(aes(group = Sample), color = "gray", alpha = 0.5) + 
  scale_color_manual(values = c("#2F8AC4","#E48725"), labels = c("Active" = "TB Disease", "Latent" = "TB Infection")) +
  facet_wrap(~Protein, scales = "free_y") +  
  theme_minimal()+
  labs(x = "Anticoagulant", y = "NPX", color = "") +
  theme(axis.text.x = element_text(size = 9,angle=0),  # Increase size of x-axis text
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size=12),
    panel.background = element_rect(fill = "white", color = NA),  
    plot.background = element_rect(fill = "white", color = NA),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )+
  stat_pvalue_manual(
    p_values, 
    label = "p_label", 
    y.position = max(input_data$NPX, na.rm = TRUE) * 1.05
  )
Fig5_scatterplot_markers
```

# Visulizing correlation between EDTA and Heparin for GSEA results {.tabset}

## 12-marker

```{r}
l=load("../Results/GSEA/12-marker.RData")
GSE_results_edta$SampleID <- gsub(".e","",GSE_results_edta$SampleID)

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

input_data <- merge(GSE_results_edta,GSE_results_TB_sweden,by="SampleID")
colnames(input_data) <- c("SampleID","Score_EDTA","Type_EDTA","Score_heparin","Type_heparin","Grp1","Grp2")

Fig5_correlationplot <- scatterplot_func2(input_data,"Score_EDTA","Score_heparin","Type_heparin","EDTA","Heparin")+
  ggtitle("12-marker signature")
Fig5_correlationplot
```

## 6-marker

```{r}
l=load("../Results/GSEA/6-marker.RData")
GSE_results_edta$SampleID <- gsub(".e","",GSE_results_edta$SampleID)

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

input_data <- merge(GSE_results_edta,GSE_results_TB_sweden,by="SampleID")
colnames(input_data) <- c("SampleID","Score_EDTA","Type_EDTA","Score_heparin","Type_heparin")

Fig5_correlationplot_B <- scatterplot_func2(input_data,"Score_EDTA","Score_heparin","Type_heparin","EDTA","Heparin")+
  ggtitle("6-marker signature")
Fig5_correlationplot_B
```

## 4-marker-1

```{r}
l=load("../Results/GSEA/4-marker-1.RData")
GSE_results_edta$SampleID <- gsub(".e","",GSE_results_edta$SampleID)

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

input_data <- merge(GSE_results_edta,GSE_results_TB_sweden,by="SampleID")
colnames(input_data) <- c("SampleID","Score_EDTA","Type_EDTA","Score_heparin","Type_heparin")

Fig5_correlationplot_E <- scatterplot_func2(input_data,"Score_EDTA","Score_heparin","Type_EDTA","EDTA","Heparin")+
  ggtitle("4-marker signature")
Fig5_correlationplot_E
```

## 4-marker-2

```{r}
l=load("../Results/GSEA/4-marker-2.RData")
GSE_results_edta$SampleID <- gsub(".e","",GSE_results_edta$SampleID)

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

input_data <- merge(GSE_results_edta,GSE_results_TB_sweden,by="SampleID")
colnames(input_data) <- c("SampleID","Score_EDTA","Type_EDTA","Score_heparin","Type_heparin")

Fig5_correlationplot_F <- scatterplot_func2(input_data,"Score_EDTA","Score_heparin","Type_EDTA","EDTA","Heparin")+
  ggtitle("4-marker signature")
Fig5_correlationplot_F
```

## 4-marker-3

```{r}
l=load("../Results/GSEA/4-marker-3.RData")
GSE_results_edta$SampleID <- gsub(".e","",GSE_results_edta$SampleID)

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

input_data <- merge(GSE_results_edta,GSE_results_TB_sweden,by="SampleID")
colnames(input_data) <- c("SampleID","Score_EDTA","Type_EDTA","Score_heparin","Type_heparin")

Fig5_correlationplot_M <- scatterplot_func2(input_data,"Score_EDTA","Score_heparin","Type_EDTA","EDTA","Heparin")+
  ggtitle("4-marker signature")
Fig5_correlationplot_M
```

# Correlation between clinical blood markers and enrichment score {.tabset}

## CRP

### Which signature? {.tabset}

#### 12-marker

```{r}
l=load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/12-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

CRP_Table <- finalTable%>%
  filter(!is.na(CRP))
CRP_Table <- CRP_Table%>% 
  rename("SampleID" = "sample_id")

CRP_Table <- CRP_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

CRP_Table <- CRP_Table%>%
  filter(Type!="Control")

correlationplot_CRP_A <- scatterplot_func2(CRP_Table,"Score","CRP","Type","Enrichment score","CRP")
```

#### 6-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/6-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

CRP_Table <- finalTable%>%
  filter(!is.na(CRP))
CRP_Table <- CRP_Table%>% 
  rename("SampleID" = "sample_id")

CRP_Table <- CRP_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

CRP_Table <- CRP_Table%>%
  filter(Type!="Control")

correlationplot_CRP_B <- scatterplot_func2(CRP_Table,"Score","CRP","Type","Enrichment score","CRP")
```

#### 4-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/4-marker-2.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

CRP_Table <- finalTable%>%
  filter(!is.na(CRP))
CRP_Table <- CRP_Table%>% 
  rename("SampleID" = "sample_id")

CRP_Table <- CRP_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

CRP_Table <- CRP_Table%>%
  filter(Type!="Control")

correlationplot_CRP_C <- scatterplot_func2(CRP_Table,"Score","CRP","Type","Enrichment score","CRP")
```


## ESR

### Which signature? {.tabset}

#### 12-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/12-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

ESR_Table <- finalTable%>%
  filter(!is.na(ESR))
ESR_Table <- ESR_Table%>% 
  rename("SampleID" = "sample_id")

ESR_Table <- ESR_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

ESR_Table <- ESR_Table%>%
  filter(Type!="Control")

correlationplot_ESR_A <- scatterplot_func2(ESR_Table,"Score","ESR","Type","Enrichment score","ESR")
```

#### 6-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/6-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

ESR_Table <- finalTable%>%
  filter(!is.na(ESR))
ESR_Table <- ESR_Table%>% 
  rename("SampleID" = "sample_id")

ESR_Table <- ESR_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

ESR_Table <- ESR_Table%>%
  filter(Type!="Control")

correlationplot_ESR_B <- scatterplot_func2(ESR_Table,"Score","ESR","Type","Enrichment score","ESR")
```

#### 4-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/4-marker-2.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

ESR_Table <- finalTable%>%
  filter(!is.na(ESR))
ESR_Table <- ESR_Table%>% 
  rename("SampleID" = "sample_id")

ESR_Table <- ESR_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

ESR_Table <- ESR_Table%>%
  filter(Type!="Control")

correlationplot_ESR_C <- scatterplot_func2(ESR_Table,"Score","ESR","Type","Enrichment score","ESR")
```

## WBC

### Which signature? {.tabset}

#### 12-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/12-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

WBC_Table <- finalTable%>%
  filter(!is.na(WBC))
WBC_Table <- WBC_Table%>% 
  rename("SampleID" = "sample_id")

WBC_Table <- WBC_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

WBC_Table <- WBC_Table%>%
  filter(Type!="Control")

correlationplot_WBC_A <- scatterplot_func2(WBC_Table,"Score","WBC","Type","Enrichment score","WBC")
```

#### 6-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/6-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

WBC_Table <- finalTable%>%
  filter(!is.na(WBC))
WBC_Table <- WBC_Table%>% 
  rename("SampleID" = "sample_id")

WBC_Table <- WBC_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

WBC_Table <- WBC_Table%>%
  filter(Type!="Control")

correlationplot_WBC_B <- scatterplot_func2(WBC_Table,"Score","WBC","Type","Enrichment score","WBC")
```

#### 4-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/4-marker-2.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

WBC_Table <- finalTable%>%
  filter(!is.na(WBC))
WBC_Table <- WBC_Table%>% 
  rename("SampleID" = "sample_id")

WBC_Table <- WBC_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

WBC_Table <- WBC_Table%>%
  filter(Type!="Control")

correlationplot_WBC_C <- scatterplot_func2(WBC_Table,"Score","WBC","Type","Enrichment score","WBC")
```

## HB

### Which signature? {.tabset}

#### 12-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/12-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

HB_Table <- finalTable%>%
  filter(!is.na(HB))
HB_Table <- HB_Table%>% 
  rename("SampleID" = "sample_id")

HB_Table <- HB_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

HB_Table <- HB_Table%>%
  filter(Type!="Control")

correlationplot_HB_A <- scatterplot_func2(HB_Table,"Score","HB","Type","Enrichment score","HB")
```

#### 6-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/6-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

HB_Table <- finalTable%>%
  filter(!is.na(HB))
HB_Table <- HB_Table%>% 
  rename("SampleID" = "sample_id")

HB_Table <- HB_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

HB_Table <- HB_Table%>%
  filter(Type!="Control")

correlationplot_HB_B <- scatterplot_func2(HB_Table,"Score","HB","Type","Enrichment score","HB")
```

#### 4-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/4-marker-2.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

HB_Table <- finalTable%>%
  filter(!is.na(HB))
HB_Table <- HB_Table%>% 
  rename("SampleID" = "sample_id")

HB_Table <- HB_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

HB_Table <- HB_Table%>%
  filter(Type!="Control")

correlationplot_HB_C <- scatterplot_func2(HB_Table,"Score","HB","Type","Enrichment score","HB")
```

## ALB

### Which signature? {.tabset}

#### 12-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/12-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

ALB_Table <- finalTable%>%
  filter(!is.na(ALB))
ALB_Table <- ALB_Table%>% 
  rename("SampleID" = "sample_id")

ALB_Table <- ALB_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

ALB_Table <- ALB_Table%>%
  filter(Type!="Control")

correlationplot_ALB_A <- scatterplot_func2(ALB_Table,"Score","ALB","Type","Enrichment score","ALB")
```

#### 6-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/6-marker.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

ALB_Table <- finalTable%>%
  filter(!is.na(ALB))
ALB_Table <- ALB_Table%>% 
  rename("SampleID" = "sample_id")

ALB_Table <- ALB_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

ALB_Table <- ALB_Table%>%
  filter(Type!="Control")

correlationplot_ALB_B <- scatterplot_func2(ALB_Table,"Score","ALB","Type","Enrichment score","ALB")
```

#### 4-marker

```{r}
load("../Results/Swedish_Clinical_Table.RData")
l=load("../Results/GSEA/4-marker-2.RData")

GSE_results_TB_sweden <- GSE_results_TB_sweden%>%
  mutate(SampleID=gsub("\\.0","",SampleID))

ALB_Table <- finalTable%>%
  filter(!is.na(ALB))
ALB_Table <- ALB_Table%>% 
  rename("SampleID" = "sample_id")

ALB_Table <- ALB_Table%>%
  left_join(GSE_results_TB_sweden,by = "SampleID")

ALB_Table <- ALB_Table%>%
  filter(Type!="Control")

correlationplot_ALB_C <- scatterplot_func2(ALB_Table,"Score","ALB","Type","Enrichment score","ALB")
```

# Save data

```{r}
save(Fig5_correlationplot_markers,Fig5_correlationplot,Fig5_correlationplot_B,Fig5_correlationplot_E,Fig5_correlationplot_F,Fig5_correlationplot_M,file="../Results/Figures/Fig5_correlationplot.RData")
save(correlationplot_CRP_A,correlationplot_CRP_B,correlationplot_CRP_C,correlationplot_ALB_A,correlationplot_ALB_B,correlationplot_ALB_C,correlationplot_HB_A,correlationplot_HB_B,correlationplot_HB_C,correlationplot_WBC_A,correlationplot_WBC_B,correlationplot_WBC_C,correlationplot_ESR_A,correlationplot_ESR_B,correlationplot_ESR_C,file="../Results/Figures/Fig7_correlationplot.RData")
```

